---
title: 
author: Kenneth Kihara
---

#Summary
The purpose of this project is to predict the way in which a person 
performed an exercise based off data taken from devices that measure 
various types of body movements. The model built predicts this with 
96% accuracy based on a 3-fold cross validation making the out of 
sample error 4%.

#Analysis
First, the dataset and appropriate packages need to be loaded.
```{r}
library(caret)
training <- read.csv("pml-training.csv")
testing <- read.csv("pml-testing.csv")

x_train <- training[,-160]
y_train <- training[,160]
x_test <- testing[,-160]
```

Decided to remove the features that contained NA values, since those 
features are mostly NA.
```{r}
x_train[x_train=="" | x_train=="#DIV/0!"] <- NA

comp <- apply(x_train, 2, function(z) all(complete.cases(z)))
x_train <- x_train[,comp]
x_train <- x_train[,-c(1:6)]
x_test <- x_test[,comp]
x_test <- x_test[,-c(1:6)]
```

Split the data to use a portion for cross validation.
```{r}
set.seed(0)
inTraining <- createDataPartition(y_train, p=.75, list=FALSE)
x_cv <- x_train[-inTraining,]
x_train <- x_train[inTraining,]
y_cv <- y_train[-inTraining]
y_train <- y_train[inTraining]
```

Train the model using the k nearest neighbor method with k = 5, since 
there are 5 classes to predict. Also, used 3-fold cross validation to 
check accuracy. The data was preprocessed to choose features using the 
PCA method and center and scale to normalize the data.
```{r}
ctrl <- trainControl(method="repeatedcv",
                     number=3,
                     repeats=5,
                     classProbs=TRUE)
tuneknn <- expand.grid(.k=5)
knnfit <- train(x=x_train, y=y_train, method="knn",
                preProcess=c("pca","center","scale"),
                trControl=ctrl, tuneGrid=tuneknn)
```

Used the partition dataset that was set aside previously to confirm 
accuracy of the model.
```{r}
knntable <- table(predict(knnfit, x_cv), y_cv)
mat <- confusionMatrix(knntable)
mat$overall
```

#Results
Accuracy was determined to be 96% making the out of sample error 4%.

Finally the test dataset was put through the model to give predictions 
on the test dataset.
```{r}
pred <- as.character(predict(knnfit, x_test))
pred
```